---
title: "clean_pupil_files"
author: "Aske Svane Qvist"
date: "15 April 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Gather Pupil Files 

In this script, all the output files from the Eye-Tracker will be cleaned up and gathered in a single csv-file. Redundant columns will be dropped and participantID and block number from the filename will be added to columns in the dataframe. These correspond to the participantID and block columns in the 'boris_cleaned' file and will thus enable merging the data. The final dataframe will be saved as 'pupil_cleaned.csv' in the folder 'gathered_files'.

## Packages
```{r}
# I use pacman to import packages
library(pacman)
p_load(tidyverse, stringi, janitor)

```

## Import pupil data
```{r}
# Save path to the csv files
path <- "../data/Pupil_Labs_raw_data/"

# Create list of filenames from the folder the path is directing to.
filenames_pupil <- list.files(path = path, pattern = ".csv")

# Create empty 'container' to which the cleaned data from each file will be appended.
data_clean <- tibble()

# Loop over all files in the folder, clean and tidy it, and combine into one great dataframe.
# The loop one time per file in the folder.
for (i in 1:length(filenames_pupil)){
  
  # Get filenames one at a time
  filename <- filenames_pupil[i]
  
  # Get participantID and block from the filename. 
  # regmatches() and regexpr() are functions working with regular expressions. The code will take the first number(s)
  # appearing in a string. If the filename is "Part19_build2.csv", '19' will be extracted.
  participantID <- as.numeric(regmatches(filename, regexpr("[0-9]*[0-9]", filename)))
  # stri_extract_last_regex() is also a function working with regular expressions. The function will extract the last
  # number in a string.  If the filename is "Part19_build2.csv", '2' will be extracted.
  block <- as.numeric(stri_extract_last_regex(filename, "\\d{1}"))
  
  # Import data from filename. The filename and the path is pasted together.
  filepath = paste(path, filename, sep = "")
  pupil <- read_csv(filepath)
  
  # Tidy up the data
  pupil <- pupil %>% 
    
    # Create ID and block columns
    mutate(
      participantID = participantID,
      block = block) %>% 
    
    # Drop unnecessary columns and only select relevant ones
    dplyr::select(2,3,4,5,6,7,9,15,16) %>% 
    # Dropped columns: 'id', 'dispersion', 'method', 'gaze_point_3d_x, 
    # 'gaze_point_3d_y', 'gaze_point_3d_z', and  'base_data'
    
    # Change duration to duration_ms as it is easier to understand
    rename(duration_ms = duration)
  
  # append data to tibble outside the loop
  # The data will be appended under the already appended data.
  if (nrow(data_clean) == 0) {
    data_clean <- pupil}
  else {
    data_clean <- rbind(data_clean, pupil)}
}


```

## Overview, write to csv and clean environment
```{r}
# Quick overview to spot potential mistakes
summary <- data_clean %>% 
  group_by(participantID, block) %>% 
  summarise(fixations = n())

# save as csv file
write_csv(data_clean,"../data/gathered_files/pupil_cleaned.csv")


# remove temporary loop objects from the environment
rm(pupil, block, filename, filenames_pupil, filepath, i, participantID, path, summary, data_clean)

```



