---
title: "changing_timestamps"
author: "Aske Svane Qvist"
date: "14 April 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Participant 4, block 1

## packages
```{r}
library(pacman)
p_load(tidyverse, rjson)


```



## Imprt json-file
```{r}
# Import timestamps from the json file
timestamps <- fromJSON(file = "../data/Pupil_Labs_raw_data/info.player/info.player_P4_B1.json")
timestamps <- as.data.frame(timestamps)

# Imoort data from pupil lab eyetracker
pupil <- read_csv("../data/Pupil_Labs_raw_data/fixations_P4_B1.csv")

# Import manually annotated information from boris
# I need to use boris raw and not cleaned.
boris <- read_csv("../data/Boris_cleaned/Part4_build1.csv")




```

## Create function changing timestamps from pupil to unix according to the document "transform timestamps into UNIX timestamps"

The timestamps in fixations.csv are Pupil Timestamps. To convert them to UNIX timestamps you could calculate the difference between start_time_system_s and start_time_synced_s from info.player.csv, and apply that difference to the timestamps in fixations.csv.
```{r}

# The function
pupil_to_unix_timestamp <- function(synced_timestamp, start_time_synced_s, start_time_system_s){
  
  # Calculate the unix timestamp
  unix_timestamp <- as.numeric(synced_timestamp) - start_time_synced_s + start_time_system_s
  
  # Return value
  return(unix_timestamp)
  
}


```



from 'Pupil' to 'Unix' timestamps starting at 0
```{r}
## Extract constants from the corresponding json file

# Pupil timestamp
start_time_synced_s <- as.numeric(timestamps[1,"start_time_synced_s"])
# Unix timestamp
start_time_system_s <- as.numeric(timestamps[1,"start_time_system_s"])


# feed the function with the pupil timestamps in the column 'start_timestamps' and the constants just extracted
pupil$start_timestamp_unix <- pupil_to_unix_timestamp(pupil$start_timestamp, start_time_synced_s, start_time_system_s)

# Subtract the unix start_time_system_s stamp.
pupil$start_timestamp_unix <- pupil$start_timestamp_unix - start_time_system_s


```

Merge the dataframes
```{r}

# Select relevant columns in pupil
pupil_df <- pupil %>% select(2,3, 6:13,15)

# Select relevant columns in boris and make appropriate column names
boris <- boris %>% 
  select(1:3, 5, 6, 8, 9, 11) %>%  
  rename(
    participantID = Part,
    block = Build,
    time = Time,
    behavior = Behavior,
    status = Status
  )

names(boris)[names(boris) == "Total length"] <- "total_length"
names(boris)[names(boris) == "Behavioral category"] <- "behavioral_category"


# Make empty data frame to be filled
data <- tibble(
  participantID = factor(), 
  block = factor(), 
  time = numeric(), 
  total_length = numeric(), 
  FPS = numeric(), 
  behavior = factor(), 
  behavioral_category = factor(),
  status = numeric(), 
  start_timestamp = numeric(),
  duration = numeric(),
  norm_pos_x = numeric(),
  norm_pos_y = numeric(),
  dispersion = numeric(),
  confidence = numeric(),
  method = numeric(),
  gaze_point_3d_x = numeric(),
  gaze_point_3d_y = numeric(),
  gaze_point_3d_z = numeric(),
  start_timestamp_unix = numeric(),
  time_dif = numeric()
)


for (fixation in 1:nrow(boris)){
  
  # Get row from boris
  row <- boris[fixation,]
  
  # get the timestamp for the fixation
  timestamp <- as.numeric(row[1,"time"])
  
  pupil_df_extract <- pupil_df %>% mutate(
    time_dif = abs(start_timestamp_unix - timestamp)
    
  )
  # Extract the row corresponding to the manually annotated fixation
  pupil_df_extract <- pupil_df_extract[pupil_df_extract$time_dif == min(pupil_df_extract$time_dif),]

  # column bind the two dfs
  combined <- cbind(row, pupil_df_extract)
  
  #Combine with premade empty dataframe
    if (nrow(data) == 0) {
      data <- combined}
    else {
      data <- rbind(data, combined)}
}


```


